---
title: "R: Basic Operations"
output: html_notebook
---

# R: Basic Operations

## Comments

Comments in R are denoted by `#`

```{r}
# This is a comment
```

There is no built-in way to do multi-line comments.

## Mathematical Operations

Mathematical operations in R are handled using the standard 4 symbols `+, -, /, and *` . Operations are calculated using PEMDAS.

### Addition

```{r}
# Addition
12+234
```

### Subtraction

```{r}
# Subtraction
456-134
```

### Multiplication

```{r}
# Multiplication
12*12
```

### Division

```{r}
# Division
180/12
```

## Variable Assignment

Variables in R are traditionally assigned using an arrow sign `<-` , made using a less than carat and a dash. It is valid to use `=` , but not recommended.

```{r}
x <- 10
x

```

Variable names can not include and spaces or symbols other than the underscore `_` . They cannot begin with numbers, but can include numbers after the first character.

```{r}
this_is_vaild_item <- 'test'
this_is_vaild_item
```

## Data Types

R has the following data types: numeric, integer, character, logical, and complex.

### Numeric

Decimal values in R are numerics. Numerics are the default data type for numbers.

```{r}
a <- 10
class(a)
```

```{r}
is.integer(a)

```

Note: Numerics are not integers. You can confirm this with both a and b above using the `is.integer()` function.

```{r}
b <- 1.3
b
```

```{r}
is.numeric(b)
```

### Integer

To create an integer in R we can use the `as.integer()` function.

```{r}
c <- as.integer(2)
is.integer(c)
```

```{r}
is.numeric((c))
```

For shorthand, we can also define an integer by writing the desired integer value, followed by a capital L.

```{r}
d <- 10L
```

```{r}
is.integer(d)
```

Integers must be whole numbers, if passed a floating value, the number will be rounded to an integer.

```{r}
e <- 10.2
as.integer(e)
```

This also works with a string.

```{r}
f <- '12.4'
as.integer(f)
```

We would use an integer data-type when we want to represent a whole integer, and don't want to accept any other variations of numbers. The numeric data-type is less strict, it allows us to switch between different number types like integers and floats dynamically.

### Character

A **character** type is used to represent strings in R. Characters can be denoted by writing within quotes.

```{r}
# Defining a character object
name <- 'John'
```

#### Concatenating character objects

Character objects can be concatenated into a single variable using the `paste` function

```{r}
first_name <- 'Babi'
last_name <- 'Ashenafi'
full_name <- paste(first_name, last_name)
full_name

```

#### Converting into character objects

We can convert other objects into character objects using the `as.character()` function

```{r}
x <- as.character(12)
x
```

```{r}
# Note that class returns the data type of our object
class(x)
```

#### Printing character types

You can print out hardcoded character objects with the `print` function

```{r}
print('I have 2 oranges')
```

You can print out formatted character objects with the `sprintf` function. You can use the following symbols to represent formatted parts of the code, each should be preceded by a `%` symbol:

-   d and i - Integer values

-   s - Character string

-   o - Octal

-   x and X - Hexadecimal

-   f - Double precision value (For fixed point representation, for example: 21.234)

-   e and E - Double precision value (For exponential representation, for example: 2.1234e+01)

```{r}
sprintf('%s ate %d cookies', 'John', 21)
```

```{r}
sprintf("%f %e", 21.234, 21.234)

```

#### Return a substring

You can get a substring from a character object using the `substr` function

```{r}

substr("How much wood could a woodchunk chunck", start <- 5, stop <- 13)

```

#### Replacing characters in a string

You can replace characters in a string using the `sub` function

```{r}
sub('much', 'littel', 'How much wood could a woodchuck chuck')
help(sub)
```

### Logical

A logical value is commonly referred to as a boolean value. It is the result of a comparison between two variables. It is represented as `TRUE` or `FALSE.`

```{r}

x <- 1
y <- 3

```

```{r}
z <- x>y
z
```

#### Logical Operators

The standard logical operations are:

-   `&` - And

-   `|` - Or

-   `!` - Not

```{r}
x <- TRUE
y <- FALSE

x&y
```

```{r}

```

```{r}

```

```{r}

```

### Complex

A complex object in R is created using the symbol `i` . The use of the i symbol is in reference to imaginary numbers.

```{r}
z <- 1+1i
z

```

```{r}
class(z)
```

## Comparison Operators

You can use the following operators to compare two variables and return a logical object:

-   `<` Less than

-   `>` Greater than

-   `<=` Less than or equal to

-   `>=` Greater than or equal to

-   `==` Is equal to

-   `!=` Is NOT equal to

```{r}
12>=12
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

## Homework

1.  Create two strings with your first and last name, respectively. Then concatenate them, and print the new value as "My name is \<your newly created concatenated variable will go here\>"

    ```{r}
    first_name <- 'Ruth'
    last_name <- 'Kidane'
    sprintf('My name is %s %s', first_name, last_name)
    ```

2.  What values can a BOOLEAN contain? BOOLEAN belongs to Logical Values which are the TRUE values and FALSE value. And R functions interpret TRUE values as 1 and FALES values as 0.
